---
title: "A4"
author: "McKenzie Kozma"
date: "9/25/2019"
output: rmarkdown::github_document
  #pdf_document: default
---

1. Using %>% operator. 
```{r}
library(tidyverse)
library(readxl)

  2019 %>% sin()
  2019 %>% cos() %>% sin()
  2019 %>% log10() %>% tan() %>% cos() %>% sin()
  2019 %>% log2()
```

2. Fixing the SEX, AGE and TRAV_SP following the steps in Assignment 2 (This time, do it on the entire
dataset instead of the sample dataset).

3. Calculate the average age and average speed of female in the accident happened in the weekend.

```{r}
c2015 <- read_xlsx("c2015.xlsx")

c2015$SEX[c2015$SEX == "Unknown"] <- "Female"
c2015$AGE[c2015$AGE == "Less than 1"] <- "0"
c2015$AGE <- as.numeric(c2015$AGE)
c2015$AGE[is.na(c2015$AGE)] <- mean(c2015$AGE, na.rm = TRUE)

c2015$TRAV_SP <- as.numeric(str_remove(c2015$TRAV_SP, "MPH"))
c2015 <- c2015[!(is.na(c2015$TRAV_SP)), ]

c2015 %>% 
  filter(SEX == "Female" & DAY_WEEK %in% c("Saturday", "Sunday")) %>% 
  summarise_at(vars(AGE, TRAV_SP), mean, na.rm = TRUE)

```

4. Use select_if and is.numeric functions to create a dataset with only numeric variables. Print out the names of all numeric variables

5. Calculate the mean of all numeric variables using select_if and summarise_all

6. We can shortcut 3 and 4 by using summarise_if: Use summarise_if to Calculate the mean of all numeric variables. (You may need to use na.rm = TRUE to ignore the NAs)

```{r}
c2015 %>% select_if(is.numeric) %>% 
  names()

c2015 %>% select_if(is.numeric) %>% 
  summarise_all(mean, na.rm = TRUE)

c2015 %>% summarise_if(is.numeric, mean, na.rm = TRUE)
```

7. Use summarise_if to calculate the median of all numeric variables.

8. Use summarise_if to calculate the standard deviation of all numeric variables. (sd function for standard deviation)

9. Use summarise_if to calculate the number of missing values for each numeric variables. Hint: Use ~sum(is.na(.))

10. Calculate the log of the average for each numeric variable.

11. You will notice that there is one NA is produced in 10. Fix this by calculating the log of the absolute
value average for each numeric variable.

```{r}
c2015 %>% summarise_if(is.numeric, median, na.rm = TRUE)

c2015 %>% summarise_if(is.numeric, sd, na.rm = TRUE)

c2015 %>% summarise_if(is.numeric, ~sum(is.na(.)))

c2015 %>% summarise_if(is.numeric, ~log(mean(.)), na.rm = TRUE)

c2015 %>% summarise_if(is.numeric, ~log(abs(mean(., na.rm = TRUE))))
```

12. Calculate the number of missing values for each categorical variables using summarise_if

13. Calculate the number of missing values for each categorical variables using summarise_all

14. Calculate the number of states in the dataset. **Hint: You can use length(table())

15. Calculate the number of uniques values for each categorical variables using summarise_if

16. Calculate the number of uniques values for each categorical variables using summarise_all

```{r}
c2015 %>% summarise_if(~!is.numeric(.), ~sum(is.na(.)))

c2015 %>% select_if(~!is.numeric(.)) %>% 
  summarise_all(~sum(is.na(.)))

c2015 %>% group_by(STATE) %>% 
  summarise() %>% 
  nrow()

c2015 %>% summarise_if(~!is.numeric(.), ~length(unique(.)))

c2015 %>% select_if(~!is.numeric(.)) %>% 
  summarise_all(~length(unique(.)))
```

17. Print out the names of all variables that have more than 30 distinct values

18. Print out the names of all categorical variables that more than 30 distinct values

19. Print out the names of all numeric variables that has the maximum values greater than 30

```{r}
c2015 %>% select_if(~length(unique(.)) > 30) %>% 
  names()

c2015 %>% select_if(~length(unique(.)) > 30 & !is.numeric(.)) %>%
  names()

c2015 %>% select_if(is.numeric) %>% 
  select_if(~max(., na.rm = TRUE) > 30) %>% 
  names()
```

20. Calculate the mean of all numeric variables that has the maximum values greater than 30 using ‘summarise_if’

21. Calculate the mean of all numeric variables that has the maximum values greater than 30 using ‘summarise_all’

```{r}
c2015 %>% select_if(is.numeric) %>% 
  summarise_if(~max(., na.rm = TRUE) > 30, mean, na.rm = TRUE)

c2015 %>% select_if(is.numeric) %>%
  select_if(~max(., na.rm = TRUE) > 30) %>% 
  summarise_all(mean, na.rm = TRUE)
```

22. Create a dataset containing variables with standard deviation greater than 10. Call this data d1

23. Centralizing a variable is subtract it by its mean. Centralize the variables of d1 using mutate_all.
Check the means of all centralized variables to confirm that they are all zeros.

24. Standarizing a variable is to subtract it to its mean and then divide by its standard deviation. Standardize the variables of d1 using mutate_all. Check the means and standard deviation of all centralized
variables to confirm that they are all zeros (for the means) and ones (for standard deviation).

```{r}
d1 <- c2015 %>% 
  select_if(is.numeric) %>% 
  select_if(~sd(., na.rm = TRUE) > 10)

d2 <- d1 %>% 
  mutate_all(~(. - mean(., na.rm = TRUE)))
d2 %>% summarise_all(mean, na.rm = TRUE)

d3 <- d2 %>% 
  mutate_all(~(./sd(., na.rm = TRUE)))
d3 %>% summarise_all(funs(mean, sd), na.rm = TRUE)
```

